Species motif participation provides unique information about species risk of extinction

### 2023-03-31 manuscript/JAE_reply.tex
starting on the reply

@@

@@

            \item Line 42-44: I am uncertain of what this sentence is trying to convey. In the sentence prior to this one, the authors suggest that trophic level and in-degree are too coarse of descriptors. Then, instead of providing insight into why or how more detailed metrics would be desirable, they mention that the consumer’s risk of extinction is dependent on its prey. It is unclear how prey-dependent extinction of consumers is related to the coarseness of the metrics used to analyze food webs. This sentence might be highlighting a key point in the article, but as it is currently written it is very difficult to understand what this sentence intends to describe.

@@

@@

@@

@@

@@

@@

            \item Results: Overall, I found the results difficult to interpret, mainly because of the number of results presented. I had issue understanding what figure 2 was showing, since the x-axis “proportion of role” was not alluded to described as an explanatory variable anywhere in the text. I wasn’t able to determine how the authors expected the probability of persistence to change across this gradient for each of the motifs, and how the results may have deviated from that expectation. This problem was mirrored in most of the figures presented.

@@

@@

@@

>>>>

% Due June 18.

@@

Proposed intro structure: 
1. What a motif is and why we care. 2. How do motifs relate to persistence? 3. Disturbance to plant communities influences consumers differently dependeing on network structure

- but that introduces persistence before we talk about disturbance ... bit circular. 
- main thrust is that we need to introduce fewer concepts, in more depth.

@@

            \item Line 42-44: I am uncertain of what this sentence is trying to convey. In the sentence prior to this one, the authors suggest that trophic level and in-degree are too coarse of descriptors. Then, instead of providing insight into why or how more detailed metrics would be desirable, they mention that the consumer's risk of extinction is dependent on its prey. It is unclear how prey-dependent extinction of consumers is related to the coarseness of the metrics used to analyze food webs. This sentence might be highlighting a key point in the article, but as it is currently written it is very difficult to understand what this sentence intends to describe.

@@

                \textbf{R:} A primary extinction is the extinction of a species that is directly disturbed (or removed, in the case of simulated species removal) while a secondary extinction is the extinction of any species not directly disturbed (as a result of the loss of its resources to primary or earlier secondary extinctions, as the Reviewer says). As we disturb primary producers and do not simulate top-down effects, only consumers can go secondarily extinction in our framework. We now define secondary extinction parenthetically in the introduction (line XX) and methods (line XX) and hope that this is clearer.

@@

                \textbf{R:} If the Reviewer is thinking of EcoPath with EcoSim, that type of model only works with quite a small food web with very well-understood interaction strengths and initial population sizes. We are not aware of any dynamical model that is fast over a large food web - certainly this was not the case when some of the authors conducted dynamical simulations of webs with 50-100 species. [[Anyone have a better answer? From the focus on major energy flows I think it's EcoSim they're thinking of...]]

@@

                \textbf{R:} [[Alyssa doesn't know how to respond to this one]]

@@

                \textbf{R:} [[Refers to: To capture a range of plausible network architectures, we simulated networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04). ]] [[Based on many collective years of experience working with empirical food webs... is that acceptable to say?]]

@@

                \textbf{R:} A focal species is simply that for which the motif role is currently being calculated ... all species are focal species in turn. Note that as these networks are simulated the species (nodes) in them do not correspond to any particular focal species or trophic guild. The vagueness is intentional as there are no characteristics of, say, invertebrates vs. vertebrates or mammals vs. reptiles incorporated in the networks. [[Another really basic misunderstanding ]][[Seems to be this line: A species’ ``motif participation role'' is the frequency with which a focal species appears in each of the motifs present in a network (Stouffer et al., 2012). Is it clearer if we say "A species’ ``motif participation role'' is the frequency with which it appears in each of the motifs present in a network"?]]

@@

                \textbf{R:} We only considered three-species motifs in this study, as stated in lines XX, XX, and XX. The fact that only four \emph{unique} three-species motifs are possible in acyclic networks is a general feature of all acyclic networks; the other 9 possible three-species motifs contain cycles of two or three species (see~\citep{Stouffer2007} for a full list of these motifs). Note that each motif can appear within the network many times with a different set of three interacting species (e.g., a network might contain the three-species-chains `hawk-mouse-oat',  `hawk-mouse-grasshopper', and `mouse-grasshopper-oat') so that the \emph{total occurrences} of all three-species motifs in a network will be much greater than four.
                [[I think it's this line: In our case, these are four-dimensional vectors since only four three-species motifs can appear in acyclic networks]]
                [[How can we clarify this line?]]

            \item Results: Overall, I found the results difficult to interpret, mainly because of the number of results presented. I had issue understanding what figure 2 was showing, since the x-axis ``proportion of role'' was not alluded to described as an explanatory variable anywhere in the text. I wasn't able to determine how the authors expected the probability of persistence to change across this gradient for each of the motifs, and how the results may have deviated from that expectation. This problem was mirrored in most of the figures presented.

                \textbf{R:} Proportion of role is the proportion of the species' motif role made up by the particular motif in each panel. [[Ummm this is kind of dumb. Will a caption rewrite fix it?]]

@@

                \textbf{R:} We thank the Reviewer for their compliments.

@@

                \textbf{R:} Using the Bayesian network framework, we simply cannot include top-down effects on persistence. However, in a parallel study (started around the same time but by chance published earlier) some of the authors of the present study did simulate both top-down and bottom-up effects in a dynamical model. The dynamical model has its own caveats (in general, dynamical models are sensative to parameter choice and, due to the time and computational resources required, a full sensitivity analysis could not be performed). However, [[something comparing the results between papers - somewhat similar?]]

                It may also reassure the Reviewer to know that in the initial studies introducing the Bayesian network framework~\citep{}, Bayesian networks captured XX\% of the extinctions identified by dynamical models run on the same set of networks. As we use the same modelling framework as in those studies, we can reasonably expect similar results.

                Finally, consider the situation from an ecological perspective. We simulate disturbance to primary producers only (and appreciate that this may not have been clear in the previous revision). While predation certainly does influence persistence, the disturbance we introduce will first act to remove food sources from some herbivores. Some fraction of these herbivores will then go extinct, reducing resources for their predators. Some fraction of these predators will also go extinct due to the reduction in their food. It is only those predators which are affected by declining resources, do not go extinct, and are able to increase predation on their remaining prey which can cause top-down extinctions. While this obviously does occur in nature, logically this is a smaller number of extinctions than the set of herbivore and predator extinctions caused by bottom-up effects. This is also reflected in the similar results from Bayesian networks and dynammical models in the studies cited above - if top-down effects were responsible for the majority of secondary extinctions we would see quite different performance.

                None of this is to discount the reality of top-down extinctions or their interest from a scientific standpoint. However, given that our interest is in bottom-up extinctions in particular and that we are up front about the absence of top-down extinctions in our model [[]]

                [[Add a comment about how strong top-down effects could change our interpretation to discussion?]]

@@

                \textbf{R:}
                [[refocus on motif participation is definitely a good idea. Though there were interaction effects with disturbance strength, so we can't leave that out.]]
